---
title: "stocksmart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stocksmart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stocksmart)
library(magrittr)
```


# Stock assessment time series data

Time series data for all federally managed stocks are bundled with the package

The [stockAssessmentData](../reference/stockAssessmentData.html) looks like this:

```{r sad, echo = FALSE}
stockAssessmentData %>%
  tibble::as_tibble()

```

Several functions are bundled with the package to aid in filtering the data by species, region, time range, metric etc. Most functions filter using the unique species ITIS code.

## Example

Problem: we want to find the latest catch data for Atlantic cod in Georges Bank from either a Benchmark assessment or a full update.

We first need to find the ITIS code for Atlantic cod. We can use the `get_species_itis` function to find this

```{r itis, echo = TRUE}
get_species_itis(stock = "Atlantic cod")

```
There are three stocks under the jurisdiction of the NEFMC, a Georges Bank, an Eastern Georges Bank, and a Gulf of Maine stock.

### Visualize data

Lets visualize all the Catch data for every assessment of the Georges Bank stock

```{r codcatch, echo = TRUE, fig.alt = "Catch data for Atlantic cod in Georges Bank"}
p <- plot_ts(itis = 164712,
             stock = "Atlantic cod - Georges Bank",
             metric = "Catch",
             printfig = FALSE)
p$plot
```

We can also plot each assessment year in its own facet

```{r codcatchfacet, echo = TRUE, fig.alt = "Catch data for Atlantic cod in Georges Bank with facets"}
p <- plot_ts(itis = 164712,
             stock = "Atlantic cod - Georges Bank",
             metric = "Catch",
             facetplot = TRUE,
             printfig = FALSE)
p$plot
```

The facet plot is particularly useful when assessment methods have changed over time and consequently the units have also. If we plot the `Abundance` instead of `Catch` we can see how the assessment data has changed over time from Metric tons prior to 2017 to kg/tow from 2017 onward

```{r codabundfacet, echo = TRUE, fig.alt = "Abundance data for Atlantic cod in Georges Bank with facets"}
p <- plot_ts(itis = 164712,
             stock = "Atlantic cod - Georges Bank",
             metric = "Abundance",
             facetplot = TRUE,
             printfig = FALSE)
p$plot
```

The [`plot_ts`](plot_ts) function returns a list of two items, a ggplot object and data frame containing the data used in the plot.

### Extract latest data

Some of the assessments visualized above may not be considered `Operational` (Analyses conducted to provide scientific advice to fishery managers with particular focus on determining stock status and recommending catch limits - from stockSMART [Data Dictionary](https://apps-st.fisheries.noaa.gov/stocksmart/StockSMART_DataDictionary.pdf)).

We can use the ITIS code to search for the most recent Catch time series data that comes from an `Operational` assessment using the `get_latest_metrics` function. A list containing two data frames are returned. 

*   A summary table containing relevant metadata including the number of years of data available and the date range.

```{r catchsummary, echo = TRUE}
cod <- get_latest_metrics(itis = 164712, metrics = "Catch")
cod$summary
```

*  A data table containing the time series data along with additional metadata

```{r catchdata, echo = TRUE}
cod$data
```

We can then filter the the data by the Georges Bank stock and plot it.

``` {r GB, echo = TRUE, fig.alt = "Catch data for Atlantic cod in Georges Bank from the latest operational assessment"}
cod$data %>%
  dplyr::filter(StockArea == "Georges Bank") %>%
  { . ->> filteredData } %>%
  ggplot2::ggplot(.) +
  ggplot2::geom_line(ggplot2::aes(x = Year, y = Value)) +
  ggplot2::ylab(filteredData %>% dplyr::distinct(Units)) +
  ggplot2::ggtitle(paste0("Assessment Year = ", filteredData %>%
                            dplyr::distinct(AssessmentYear)))
```
